---
title: "DS4B Data Wrangling Recap"
author: "Paul Apivat Hanvongse"
date: "6/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Select & Arrange

I had previously selected by column name, never by numeric vector; and I had never used select_helpers (starts_with, contains, matches, num_range). I had never re-arranged columns using select `select(col3:col5, everything())`. `Select_if()` also comes in handy as well as the `is.integer`, `is.double` data types. 

Knowing data types allow for efficient selection. 

I've used arrange() and desc(), so I don't go into too much detail here. 

```{r USArrests, echo=TRUE, results='hide'}
library(tidyverse)

# select first three columns using numeric vector
USArrests %>%
    select(1:3)

# re-arrange columns using select; place UrbanPop first
USArrests %>%
    select(UrbanPop, everything())

# select if integer 
USArrests %>%
    select_if(is.integer)

USArrests %>%
    select_if(~ !is.double(.))

# select if double
USArrests %>%
    select_if(is.double)

USArrests %>%
    select_if(~ !is.integer(.))

# swithing to IRIS dataframe
# select helpers: starts_with, contains
iris %>%
    select(starts_with("petal"))

iris %>%
    select(contains("length")) %>% 
    head()

# using pull() instead of select() to pull out contents of a column
iris %>%
    pull(Sepal.Width) %>%
    head()

iris %>% 
    select_if(is.factor) %>%
    head()

# arrange Species in descending order by Petal.Length
iris %>%
    select(Petal.Length, Species) %>%
    arrange(desc(Petal.Length))

```

## Group By and Summarize

You can also embed plots, for example:

```{r irs, echo=TRUE, results='hide'}
# describe Species by Petal Length
iris %>%
    group_by(Species) %>%
    summarize(
        count = n(),
        mean = mean(Petal.Length),
        median = median(Petal.Length),
        sd = sd(Petal.Length),
        min = min(Petal.Length),
        max = max(Petal.Length)
    )

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
